Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-02-05 article.tex
affiliations

\begin{lemma}
\label{lemma:MatchStripeToPermutation}
Given two permutations $\ptext$,$\pmotif$ avoiding (231,213), If $\stripew{\pmotif}$ is a motif of stripe $\stripew{\ptext}$ by a subsequence $s$, then $\pmotif$ is a motif of $\ptext$ by subsequence $s$.
\end{lemma}

\begin{proof} 
We prove the lemma by induction on the size of the motif.The base case is a motif of size 2. Suppose that we have the motif $(12)$, thus the stripe of the motif is $\ustep$. If there is a match in stripe, then there exists two (contiguous) increasing elements in the text, which is a match in permutation. Same goes for the permutation $(21)$. Now, suppose that the lemma is true for every motif in $\Av{n}{(231,213)}$. Let $\pmotif$ be a motif in $\Av{n+1}{(231,213)}$, and $\ptext$ a permutation avoiding $(213,231)$. If there is a match of $\stripew{\pmotif}$ in $\stripew{\ptext}$ by subsequence $s$, then there is a match of $\stripew{\pmotif}[2:]$ in $\stripew{\ptext}[2:]$ by subsequence $s[2:]$, and by induction there is a match of $\pmotif[2:]$ in $\ptext[2:]$ by subsequence $s[2:]$. Moreover if the first element the stripe of the motif is ascend (descent), and by lemma \ref{lemma:first element is 1 or n} the minimal (maximal) element, and remark that a subsequence of a permutation avoiding (231,213) is also avoiding (231,213), thus the first element of the subsequence $s$ is either the minimal (maximal) element which complete our proof.
\end{proof}

\begin{proposition}
Let $\sigma \in \AV_k(213,231)$ and $\pi \in \AV_k(213,231)$.
$\sigma$ is a motif of $\pi$ if and only if 
$s_\sigma$ is a subsequence of $s_\pi$.
\end{proposition}

		% without lost of generality we can suppose that $x<y$,
% 		Let see all the case possible that can make this 
% 		hypothesis possible :
% 		\begin{itemize}		
% 			\item if $\pmotif(x)<\pmotif(y)$, 
% 			$\stripe{\pmotif}{x}=\ustep$ 
% 			and $\stripe{\pmotif}{y}=\ustep$
% 			(if x and y are increasing in $\pmotif$),
% 			then $\stripe{\ptext}{\varphi(x)} =\ustep$ and $\stripe{\ptext}{\varphi(y)}=\ustep$,
% 			(the match are also increasing).
% 			As $x<y$ and as $\varphi$ is an increasing function, 
% 			$\varphi(x) < \varphi(y)$.
% 			And thus, by \textbf{Lemma \ref{lemma:upup}},
% 			$\ptext(\varphi(x)) < \ptext(\varphi(y))$ .
% 			Which contradict our hypothesis,
% 			so this case is not possible.

\begin{proof}
$\Rightarrow$ Suppose that $\pmotif$ is a motif of $\ptext$ by subsequence $s$. By definition $s$ normalized is equal to
$\pmotif$ so they have the same stripe.\\ 
$\Leftarrow$ This is proved by lemma \ref{lemma:MatchStripeToPermutation}.
\end{proof}

\begin{proposition}
Let $\pmotif$ and $\ptext$
be two permutations avoiding (231,213), 
we can decide in linear-time if $\pmotif$ 
appear in $\ptext$.	
\end{proposition}

>>>>

\begin{lemma}
\label{lemma:MatchStripeToPermutation}
	Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations.
	Then $\pi$ contains $\sigma$ if and only if 
	$s_\sigma$ is a subsequence of $s_\pi$.
\end{lemma}

\begin{proof} 
	The forward direction is obvious.
	We prove the backward direction by induction on the size of the motif
	$\sigma$.
	The base case is a motif of size 2. Suppose that we have the motif $(12)$, thus the stripe of the motif is $\ustep$. If there is a match in stripe, then there exists two (contiguous) increasing elements in the text, which is a match in permutation. Same goes for the permutation $(21)$. Now, suppose that the lemma is true for every motif in $\Av{n}{(231,213)}$. Let $\pmotif$ be a motif in $\Av{n+1}{(231,213)}$, and $\ptext$ a permutation avoiding $(213,231)$. If there is a match of $\stripew{\pmotif}$ in $\stripew{\ptext}$ by subsequence $s$, then there is a match of $\stripew{\pmotif}[2:]$ in $\stripew{\ptext}[2:]$ by subsequence $s[2:]$, and by induction there is a match of $\pmotif[2:]$ in $\ptext[2:]$ by subsequence $s[2:]$. Moreover if the first element the stripe of the motif is ascend (descent), and by lemma \ref{lemma:first element is 1 or n} the minimal (maximal) element, and remark that a subsequence of a permutation avoiding (231,213) is also avoiding (231,213), thus the first element of the subsequence $s$ is either the minimal (maximal) element which complete our proof.
\qed
\end{proof}

We are now ready to solve the pattern containment in case
both $\pi$ and $\sigma$ are $(213, 231)$-avoiding.

\begin{proposition}
	Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations.
	One can decide whether $\pi$ contains $\sigma$ in linear-time.
\end{proposition}
